---
title: "Untitled"
author: "JIhyun Lee"
date: "12/5/2020"
output: html_document
---
#### Bagging

Bagging uses the bootstrap technique to reduce the variance by taking repeated training samples and average over.

#### Bagging

```{r}
set.seed(201205)

bag<- randomForest(quality~., data = data_train, mtry = 12, importance=TRUE)

plot(bag, main = "Out-of-bag Error")
```


```{r}
## Test set
pred_y_bag <- predict(bag, newdata = data_tests)

plot(data_tests$quality, pred_y, pch = 16, xlab = 'observed', ylab = 'predicted')
abline(0, 1, lwd = 2, col = 'red')
```

```{r}
err_bag <- (pred_y_bag - data_tests$quality)^2

# MSE
mean(err_bag)
```


```{r}
# Importance
# plot(importance(bag), xlab = 'variables', ylab = 'importance', 
#      xaxt = 'n', type = 'h', lwd = 3)
# axis(1, at = 1:length(importance(bag)), labels = rownames(importance(bag))) ## Not sure why axis is not working

varImpPlot(bag)
```



`%IncMSE` indicates the mean decrease of accuracy in prediction on the out of bag samples when a given variable is excluded from the model. As the larger decrease, the target variable is important in the model.
`IncNodePurity` measures the total decrease in training RSS.
`Alcohol` and `Volatile acidity` seem to be important variables.


Pruning alternatives:
```{r message = F}
library(tree)

##### Pruning ?
# fit 
tree <- tree(quality~., data)
summary(tree)


## pruning the tree
cv_tree <- cv.tree(tree)
plot(cv_tree$size ,cv_tree$dev, type="b")

## pruning
prune_tree <- prune.tree(tree, best = 6)
plot(prune_tree)
text(prune_tree ,pretty=0)

# Another package
prune_tree2 <- prune.rpart(best_tree, cp = .01)

rpart.plot(prune_tree2, clip.right.labs = FALSE, under = TRUE, digits = 4) # same as before.

```